---
title: "Phage_pilot"
author: "John McMullen"
date: "11/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Initial phage pilot experiment
-Plants were seeded into three different soil types (SV = 1:1 parts sand and vermiculite; VTP = 1:1:1 parts vermiculite, turface clay, and ground peat moss; MMT = 1:2 parts metro mix 820 and turface clay) with approx. three clover seeds. Date: 22 Sept 21
-Plants were heavily waters twice daily for two weeks, at which point, plants were culled except 1 (tried to pick similar plant sizes across pots), and 500 ul 10^7 CFU ml^-1 R. leguminosarum 461 was added to base of the plant as well as 500 ul 10^6 PFU ml^-1 rhizophage depending on the microbial addition treatment. 500 ul PBS was added to aposymbiotic plants. From this point on, plants were misted once daily. Date: 7 Oct 21
-Plants were harvested for microbial and plant fitness metrics. Plants were not watered morning of harvest day. Date: 3 Nov 21

###################

Data
```{r}
library(ggplot2) #draw figures
library(emmeans) #post hoc comparisons
library(car) #Anova function
library(dplyr) #produce mean and se for plant biomass

# dat = read.csv("~/Dropbox/docs/Lennon_Lab/rhizophage/data/phage_pilot.csv")
dat = read.csv("data/phage_pilot.csv")

#convert columns in character to factors
for(i in 1:ncol(dat)) {
  ifelse(class(dat[,i])=="character",dat[,i]<-as.factor(dat[,i]),0)
}
```


#Chlorophyll content 
```{r}
#analysis is weighted by number of leaves measured
chl_w = lm(chlor_content ~soil_type*micro_trt, weights = leaf_chlor_measure,data=dat)
summary(chl_w)

#assess normality
chl_w.res.lm = residuals(chl_w)
summary(chl_w.res.lm)
qqPlot(chl_w.res.lm)
hist(chl_w.res.lm)
boxplot(chl_w.res.lm)
shapiro.test(chl_w.res.lm)

#assess homoscedasticity
plot(fitted(chl_w),chl_w.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(chl_w),chl_w.res.lm));
leveneTest(chl_w.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(chl_w)

chl_w_post = emmeans(chl_w, ~ micro_trt|soil_type)
pairs(chl_w_post)
cld(chl_w_post)



#######################

#analysis is not weighted - for comparison, but it makes sense to keep leaves measured in the model. The weights weaken the interaction effect.
chl = lm(chlor_content ~soil_type*micro_trt,data=dat)
summary(chl)

#assess normality
chl.res.lm = residuals(chl)
summary(chl.res.lm)
qqPlot(chl.res.lm)
hist(chl.res.lm)
boxplot(chl.res.lm)
shapiro.test(chl.res.lm)

#assess homoscedasticity
plot(fitted(chl),chl.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(chl),chl.res.lm));
leveneTest(chl.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(chl)

chl_post = emmeans(chl, ~ micro_trt|soil_type)
pairs(chl_post)
cld(chl_post)

#########

#Graph of the weighted model
chl_gg = as.data.frame(chl_w_post)

ggplot(chl_gg, aes(micro_trt, emmean)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab('Chlorophyll [SPAD units]\n(mean \U00B1 SE, n = 3)') +
  scale_y_continuous(limits = c(0,50),breaks=seq(0,50,by=10))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))
```


#Chlorophyll content converted
```{r}
#analysis is weighted by number of leaves measured; formula from ling et al 2011 photosynthesis research - tried just for funzies. Seems like if we can't get differences in the future, we should make a curve since this is a non linear relationship

cnvt = (0.0007*((dat$chlor_content)^2)) + (0.023*dat$chlor_content) + 0.0544
chl_w = lm(cnvt ~soil_type+micro_trt, weights = leaf_chlor_measure,data=dat)
summary(chl_w)

#assess normality
chl_w.res.lm = residuals(chl_w)
summary(chl_w.res.lm)
qqPlot(chl_w.res.lm)
hist(chl_w.res.lm)
boxplot(chl_w.res.lm)
shapiro.test(chl_w.res.lm)

#assess homoscedasticity
plot(fitted(chl_w),chl_w.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(chl_w),chl_w.res.lm));
leveneTest(chl_w.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(chl_w)

chl_w_post = emmeans(chl_w, ~ micro_trt|soil_type)
pairs(chl_w_post)
cld(chl_w_post)



#######################

#Graph of the weighted model
chl_gg = as.data.frame(chl_w_post)

ggplot(chl_gg, aes(micro_trt, emmean)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab('Chlorophyll content [nmol mg^-1]\n(mean \U00B1 SE, n = 3)') +
  #scale_y_continuous(limits = c(0,50),breaks=seq(0,50,by=10))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))
```


#Nodule biomass
```{r}
#analysis was weighted to the number of nodules (adding slightly dampens the soil effect), had to remove nodules larger than 1 mg since this was a measuring issue. interaction dropped, p = 0.52876
nod_w = lm(per_nod_mg ~soil_type+micro_trt,weights = number_nodules, data=subset(dat,per_nod_mg<1))
summary(nod_w)

#assess normality
nod_w.res.lm = residuals(nod_w)
summary(nod_w.res.lm)
qqPlot(nod_w.res.lm)
hist(nod_w.res.lm)
boxplot(nod_w.res.lm)
shapiro.test(nod_w.res.lm)

#assess homoscedasticity
plot(fitted(nod_w),nod_w.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(nod_w),nod_w.res.lm));
leveneTest(nod_w.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(nod_w)

nod_w_post = emmeans(nod_w, ~ micro_trt|soil_type)
pairs(nod_w_post)

#######################
#Graph of model
nod_gg = as.data.frame(nod_w_post)

ggplot(nod_gg, aes(micro_trt, emmean)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=emmean-SE,ymax=emmean+SE),width=0.2) +
  ylab(expression(atop('Biomass [mg '~nodule^-1~"]", paste('(mean \U00B1 SE, n = 3)')))) +
  scale_y_continuous(limits = c(-0.01,0.6),breaks=seq(0,0.6,by=0.2))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))

```


#Nodule number
```{r}
hist(dat$number_nodules)
hist(subset(dat,micro_trt != "apo")$number_nodules)

# dropped interaction p = 0.61893
nodn_p = glm(number_nodules ~soil_type+micro_trt,family=quasipoisson(link = "log"),data=subset(dat,micro_trt != "Apo"))
summary(nodn_p)

#1 - pchisq(succ.mdl$deviance, succ.mdl$df.resid)
1-pchisq(35.384,13) # 0.0007389516
35.384/13 #2.7

Anova(nodn_p,test.statistic = "Wald")

nodn_post_p = emmeans(nodn_p, ~ micro_trt|soil_type, type="response")
pairs(nodn_post_p)

#Graph of model
nodn_gg = as.data.frame(nodn_post_p)

ggplot(nodn_gg, aes(micro_trt, rate)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),width=0.2) +
  ylab('Number of nodules\n(mean \U00B1 CI, n = 3)') +
  scale_y_continuous(limits = c(0,80),breaks=seq(0,80,by=20))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))
```

#Leaf number
```{r}
hist(dat$leaf_count)

#dropped interaction p = 0.90857
leafn_p = glm(leaf_count ~soil_type+micro_trt,family=poisson(link = "log"),data=dat)
summary(leafn_p)

#1 - pchisq(succ.mdl$deviance, succ.mdl$df.resid)
1-pchisq(3.6886,21) # 0.9999902
3.6886/21 #0.1756476

Anova(leafn_p,test.statistic = "Wald")

leafn_post_p = emmeans(leafn_p, ~ micro_trt|soil_type,type="response")
pairs(leafn_post_p)

#Graph of the weighted model
leafn_gg = as.data.frame(leafn_post_p)

ggplot(leafn_gg, aes(micro_trt, rate)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=asymp.LCL,ymax=asymp.UCL),width=0.2) +
  ylab('Number of leaves\n(mean \U00B1 CI, n = 3)') +
  scale_y_continuous(limits = c(0,10),breaks=seq(0,10,by=2))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))
```

#Above ground biomass
```{r}
#model for plant biomass - log transformed to normalize residuals.
plant_w = lm(log(above_biomass_mg) ~soil_type*micro_trt,data=dat)
summary(plant_w)

#assess normality
plant_w.res.lm = residuals(plant_w)
summary(plant_w.res.lm)
qqPlot(plant_w.res.lm)
hist(plant_w.res.lm)
boxplot(plant_w.res.lm)
shapiro.test(plant_w.res.lm)

#assess homoscedasticity
plot(fitted(plant_w),plant_w.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(plant_w),plant_w.res.lm));
leveneTest(plant_w.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(plant_w)

plant_w_post = emmeans(plant_w, ~ micro_trt|soil_type,type="response")
pairs(plant_w_post)

#######################
#Graph of model
plant_gg = dat %>% group_by(micro_trt,soil_type) %>% summarise(mean = mean(above_biomass_mg,na.rm=TRUE),SE = sd(above_biomass_mg,na.rm = TRUE)/sqrt(length(na.omit(above_biomass_mg))))

ggplot(plant_gg, aes(micro_trt, mean)) +
  geom_point(size=5) +
  geom_errorbar(aes(ymin=mean-SE,ymax=mean+SE),width=0.2) +
  ylab(expression(atop('Biomass [mg'~plant^-1~"]",paste('(mean \U00B1 SE, n = 3)')))) +
  scale_y_continuous(limits = c(0,100),breaks=seq(0,100,by=20))+
  theme_bw(base_size = 30) +
  facet_grid(~soil_type,scale="free")+
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))

```


```{r}
#associate plant biomass in relation to microbial treatment, soil, and number of nodules (proxy for symbiont fitness). Tried interacting with nodule number, but the data looked a little goofy for what pulled out differences due to the fact that there is a low number of reps (I think).

cor_biomass = lm((above_biomass_mg) ~ (number_nodules)*micro_trt*soil_type,data=subset(dat,micro_trt!="Apo"))
summary(cor_biomass)
AIC(cor_biomass)

#assess normality
cor_biomass.res.lm = residuals(cor_biomass)
summary(cor_biomass.res.lm)
qqPlot(cor_biomass.res.lm)
hist(cor_biomass.res.lm)
boxplot(cor_biomass.res.lm)
shapiro.test(cor_biomass.res.lm)

#assess homoscedasticity
plot(fitted(cor_biomass),cor_biomass.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(cor_biomass),cor_biomass.res.lm));
leveneTest(cor_biomass.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(cor_biomass,type='III')

emtrends(cor_biomass, pairwise ~ micro_trt|soil_type, var = "number_nodules")

mmtavg = mean(c(dat[dat$micro_trt!="Apo"&dat$soil_type=='SV',12]))
emmeans(cor_biomass,pairwise~micro_trt|soil_type,at=list(number_nodules=mmtavg))

ggplot(subset(dat,micro_trt!="Apo"),aes(x=number_nodules,y=above_biomass_mg,color=micro_trt)) +
  geom_point(size=5,aes(shape=soil_type)) +
  ylab(expression("Plant biomass [mg]")) +
  xlab(expression("Number of nodules"))+
  geom_smooth(method="lm") +
  #annotate("text", x = 10, y = 95, label = "r = 0.95 \n p < 0.0001",size=8)+
  theme_bw(base_size = 30) +
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))

fit_cor = subset(dat,micro_trt!="Apo")
cor.test(fit_cor$number_nodules, fit_cor$above_biomass_mg) #0.9516598 
```

```{r}
#associate plant biomass in relation to microbial treatment, soil, and nodule biomass (proxy for symbiont fitness) - can also do with per nodule biomass but only positive association for MMT. Not interaction with MMT nodule biomass and trt, removed

cor_b_biomass = lm(log(above_biomass_mg) ~ per_nod_mg+micro_trt,data=subset(dat,micro_trt!="Apo"&soil_type=="MMT"))
summary(cor_b_biomass)
AIC(cor_b_biomass)

#assess normality
cor_b_biomass.res.lm = residuals(cor_b_biomass)
summary(cor_b_biomass.res.lm)
qqPlot(cor_b_biomass.res.lm)
hist(cor_b_biomass.res.lm)
boxplot(cor_b_biomass.res.lm)
shapiro.test(cor_b_biomass.res.lm)

#assess homoscedasticity
plot(fitted(cor_b_biomass),cor_b_biomass.res.lm);
abline(h=0,lty=2);
lines(smooth.spline(fitted(cor_b_biomass),cor_b_biomass.res.lm));
leveneTest(cor_b_biomass.res.lm ~ soil_type*micro_trt, data = dat)
#looks okay

Anova(cor_b_biomass)

ggplot(subset(dat,micro_trt!="Apo"),aes(x=per_nod_mg,y=above_biomass_mg,color=micro_trt,shape=soil_type)) +
  geom_point(size=5) +
    theme_bw(base_size = 30) +
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))

ggplot(subset(dat,micro_trt!="Apo"&soil_type=="MMT"),aes(x=per_nod_mg,y=above_biomass_mg,color=micro_trt,shape=soil_type)) +
  geom_point(size=5) +
    theme_bw(base_size = 30) +
  theme(aspect.ratio=1,panel.grid = element_blank(),
        panel.border = element_rect(size = 1, colour = "black"),
        legend.title = element_blank(),axis.title.x = element_blank(),axis.text.x=element_text(size=rel(0.9)),axis.text.y=element_text(size=rel(0.9)),axis.title.y=element_text(size=rel(0.9)))

fit_cor = subset(dat,micro_trt!="Apo")
cor.test(fit_cor$per_nod_mg, fit_cor$above_biomass_mg) #0.9516598 
```

